---
title: "Brain Stroke Prediction"
author: "Team Sage"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---


Brain Stroke:

A stroke is a medical condition in which poor blood flow to the brain causes cell death. According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

Dataset Link: https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset
```{r}
library(ggplot2)
library(corrplot)
library(ROSE)
```

```{r}
stroke_df = read.csv("healthcare-dataset-stroke-data.csv")
stroke_df = subset(stroke_df, select = -c(id))
head(stroke_df)
# Getting the summary of the dataframe
summary(stroke_df)
# Looking at the structure of the Dataframe
str(stroke_df)
```
```{r}
# changing the variables to categorical variables and numerical variables
stroke_df$gender = as.factor(stroke_df$gender)
stroke_df$hypertension = as.factor(stroke_df$hypertension)
stroke_df$heart_disease = as.factor(stroke_df$heart_disease)
stroke_df$ever_married = as.factor(stroke_df$ever_married)
stroke_df$bmi = as.numeric(stroke_df$bmi)
stroke_df$stroke = as.factor(stroke_df$stroke)
str(stroke_df)
# Finding if there are any NA's in the dataset
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
summary(stroke_df)
```
```{r}
# There are 201 NA's in BMI. Replacing those variables with mean of BMI
stroke_df$bmi[is.na(stroke_df$bmi)] = mean(stroke_df$bmi,na.rm = TRUE)
# Finding if there are any NA's after imputation of null values
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
```

```{r}
ggplot(stroke_df, aes(x=stroke,fill=stroke))+geom_bar()+ggtitle("Distribution of Target varibale(Stroke)")+ theme_bw() +  theme()+ xlab("Stroke") + ylab("Count of people")
```


Since the Data is imbalanced we are using sampling techniques such as under sampling and over sampling. In this case we used both under and over sampling 
```{r}
# we are first splitting the data into test and train 
set.seed(123)
ind <- sample(2, nrow(stroke_df), replace = TRUE, prob = c(0.7, 0.3))
train <- stroke_df[ind==1,]
test <- stroke_df[ind==2,]
# In the train data we are using under and oversampling to address the class imbalance problem
both <- ovun.sample(stroke~., data=train, method = "both",
                    p = 0.5,
                    seed = 222,
                    N = 5110)$data
r<-table(both$stroke)
paste("Seeing if the class imabalance is fixed")
r
# copying this data frame to a new dataframe
stroke_df_train_sample = both
str(stroke_df_train_sample)
# Looking at the Information in the dataset after sampling 
summary(both)
```
```{r}
ggplot(stroke_df_train_sample, aes(x=stroke,fill=stroke))+geom_bar()+ggtitle("Distribution of Target varibale after sampling")+ theme_bw() +  theme()+ xlab("Stroke") + ylab("Count of people")
```

```{r}

ggplot(stroke_df_train_sample, aes(x=smoking_status,fill=stroke))+geom_bar(position = "dodge")+ggtitle("People segregated based on their smoking habbits")+ theme_bw() +  theme()+ xlab("Smoking status") + ylab("Number of people")
```

```{r}
# Count of Males and Females in the Dataset
ggplot(stroke_df_train_sample, aes(x=gender,fill=gender))+geom_bar()+ggtitle("Count of Male and Female in the Dataset")+ theme_bw() +
  theme()+ xlab("Gender") + ylab("Count of people")
# Seeing if work type affects people
ggplot(stroke_df_train_sample,aes(x=work_type, fill=stroke))+geom_bar(position = "dodge")+ggtitle("Work Type Vs Stroke")+ theme_bw() +
  theme()+ xlab("Work Type") + ylab("Count of people who had a stroke")
# Seeing if Smoking status causes stroke
ggplot(stroke_df_train_sample,aes(x=work_type, fill= smoking_status))+geom_bar(position = "dodge")+ggtitle("Smoking habbits in different work type")+ theme_bw() + theme()+ xlab("Work Type") + ylab("Count of people ")
```

```{r}
# Bmi distribution
ggplot( stroke_df_train_sample,aes(x=bmi)) + geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)+ theme_bw() + theme()+ ggtitle("BMI distribution")
# BMI vs stroke 
ggplot(stroke_df_train_sample, aes(x=bmi, fill=stroke)) + geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity')+ theme_bw() + theme()

```
```{r}
# checking if any variables are related
stroke_numeric = subset(stroke_df_train_sample, select = c(age,avg_glucose_level,bmi))
corr_matrix = cor(stroke_numeric)
corrplot(corr_matrix, method = "number", type = "upper")

```

```{r}
stroke_1 = subset(stroke_df_train_sample, stroke == 1)
stroke_0 = subset(stroke_df_train_sample, stroke == 0)
# how bmi and age is stroke in people 
ggplot(stroke_1, aes(x=bmi, y=age, color=stroke))+geom_point(size=3)
ggplot(stroke_0, aes(x=bmi, y=age, color=stroke))+geom_point(size=3)
ggplot(stroke_1, aes(x=bmi, y=age, color=stroke))+geom_boxplot()
ggplot(stroke_0, aes(x=bmi, y=age, color=stroke))+geom_boxplot()
ggplot(stroke_1, aes(x=age, y=bmi, color=stroke))+geom_boxplot()
ggplot(stroke_0, aes(x=age, y=bmi, color=stroke))+geom_boxplot()
# we can see 25 % to 75 % who get stroke are from age 58 - 78 . the higher the age the chances of getting a stroke is more
# we can see become who have bmi from 27.5 to 32.5 have higher chances of stroke
```
```{r}
ctest = chisq.test(as.factor(stroke_df_train_sample$smoking_status), stroke_df_train_sample$stroke)
ctest
```

The pvalue of this test is <0.05 so we fail to accept the null hypothesis. That is these two variables smoking status and stroke are related.

```{r}
ctest1 = chisq.test(as.factor(stroke_df_train_sample$work_type), stroke_df_train_sample$stroke)
ctest1
```
The pvalue of this test is <0.05 so we fail to accept the null hypothesis. That is these two variables work type and stroke are related.


