---
title: "Brain Stroke Prediction"
author: "Team Sage"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---


Brain Stroke:

A stroke is a medical condition in which poor blood flow to the brain causes cell death. According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

Dataset Link: https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset
```{r}
library(ggplot2)
library(gridExtra)
library(corrplot)
library(ROSE)
library(dplyr)
```

```{r}

stroke_df = read.csv("healthcare-dataset-stroke-data.csv")
stroke_df = subset(stroke_df, select = -c(id))
head(stroke_df)

# Getting the summary of the dataframe
summary(stroke_df)


# Looking at the structure of the Dataframe
str(stroke_df)

```
```{r}
# changing the variables to categorical variables and numerical variables
stroke_df$gender = as.factor(stroke_df$gender)
stroke_df$hypertension = as.factor(stroke_df$hypertension)
stroke_df$heart_disease = as.factor(stroke_df$heart_disease)
stroke_df$ever_married = as.factor(stroke_df$ever_married)
stroke_df$bmi = as.numeric(stroke_df$bmi)
stroke_df$stroke = as.factor(stroke_df$stroke)
str(stroke_df)

# Finding if there are any NA's in the dataset
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
summary(stroke_df)

```
```{r}
# There are 201 NA's in BMI. Replacing those variables with mean of BMI
stroke_df$bmi[is.na(stroke_df$bmi)] = mean(stroke_df$bmi,na.rm = TRUE)
# Finding if there are any NA's after imputation of null values
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
```

```{r}

ggplot(stroke_df, aes(x=stroke,fill=stroke))+geom_bar()+ggtitle("Distribution of Target varibale(Stroke)")+ theme_bw() +  theme()+ xlab("Stroke") + ylab("Count of people")
```


Since the Data is imbalanced we are using sampling techniques such as under sampling and over sampling. In this case we used both under and over sampling 
```{r}
# we are first splitting the data into test and train 
set.seed(123)
ind <- sample(2, nrow(stroke_df), replace = TRUE, prob = c(0.7, 0.3))
train <- stroke_df[ind==1,]
test <- stroke_df[ind==2,]


# In the train data we are using under and oversampling to address the class imbalance problem

both <- ovun.sample(stroke~., data=train, method = "both",
                    p = 0.5,
                    seed = 222,
                    N = 5110)$data
r<-table(both$stroke)
paste("Seeing if the class imabalance is fixed")
r
# copying this data frame to a new dataframe
stroke_df_train_sample = both
str(stroke_df_train_sample)
# Looking at the Information in the dataset after sampling 
summary(both)

```

```{r}

ggplot(stroke_df_train_sample, aes(x=stroke,fill=stroke))+geom_bar()+ggtitle("Distribution of Target varibale after sampling")+ theme_bw() +  theme()+ xlab("Stroke") + ylab("Count of people")

```

```{r}
ggplot(stroke_df, aes(x = age)) + 
    geom_histogram(fill = 'blue', col = 'black', binwidth = 7, size=1)

ggplot(stroke_df, aes(x = age, fill = stroke))+
  geom_density(alpha = 0.3)
```

```{r}
ggplot(data = stroke_df, aes(x=ever_married, y=age, fill=ever_married)) + geom_boxplot()+
  labs(title="Age distribution by Marital status", x="ever_married", y="age")
```

```{r}
ggplot(data = stroke_df, aes(x=as.character(hypertension), y=age, fill=hypertension)) +
    geom_boxplot() +
    labs(title="Age distribution by hypertension", x="hypertension", y="age")

# not needed
ggplot(data = stroke_df, aes(x=as.character(hypertension), y=avg_glucose_level, fill=hypertension)) + geom_boxplot() +
    labs(title="Average Glucose Level distribution by hypertension", x="hypertension", y="avg_glucose_level")
```

```{r}
stroke_df %>%
  filter(age >= 50, age <= 65) %>%
  select(hypertension, stroke) %>%
  table()
```

```{r}
# Checking Test of Independence between hypertension and Target Variable Stroke
#alpha =0.05
ht_ch<-table(stroke_df$hypertension, stroke_df$stroke) 
print(ht_ch)
'When hypertension is 0 (i.e., no hypertension) then there are 4429 people with no stroke and 183 with stroke, and when hypertension is 1 (i.e., hypertension is present) then there are 432 people who had no stroke, and 66 people with stroke.'

chisq.test(stroke_df$hypertension, stroke_df$stroke)
```
p-value < 0.05, so we can say that there is statistically significant difference in stroke frequency between groups of people in age of 50-65 with and without hypertension.
```{r}
ggplot(data = stroke_df, aes(x=as.character(heart_disease ), y=age, fill=heart_disease)) +
    geom_boxplot() +
    labs(title="Age distribution by heart_disease ", x="heart_disease ", y="age")

# not needed
ggplot(data = stroke_df, aes(x=as.character(heart_disease ), y=avg_glucose_level, fill=heart_disease)) + geom_boxplot() +
    labs(title="Average Glucose Level distribution by heart_disease ", x="heart_disease ", y="avg_glucose_level")
```

```{r}
stroke_df %>%
  filter(age >= 50, age <= 65) %>%
  select(heart_disease, stroke) %>%
  table()
```

```{r}
# Checking Test of Independence between heart_disease and Target Variable Stroke
#alpha =0.05
hd_ch<-table(stroke_df$heart_disease, stroke_df$stroke)
print(hd_ch)
"There are a total of 4632 people who didn't suffer a stroke before and there are 202 people who did not have a stroke before but did have stroke later. On the other hand, there are 229 people who previously suffered with a stroke but not later, and 47 who suffered with a stroke before and later as well."

chisq.test(stroke_df$heart_disease, stroke_df$stroke)
```
p-value < 0.05, so we can say that there is statistically significant difference in stroke frequency between groups of people in age of 50-65 with and without heart disease.