---
title: "BRAIN STROKE PREDICTION"
team:  "SAGE"

---

Brain Stroke:

A stroke is a medical condition in which poor blood flow to the brain causes cell death. According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

Dataset Link: https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset


```{r}

stroke_df = read.csv("healthcare-dataset-stroke-data.csv")
stroke_df = subset(stroke_df, select = -c(id))
head(stroke_df)

# Getting the summary of the dataframe
summary(stroke_df)


# Looking at the structure of the Dataframe
str(stroke_df)

```
```{r}
# changing the variables to categorical variables and numerical variables
stroke_df$gender <- as.factor(stroke_df$gender)
stroke_df$hypertension <- as.factor(stroke_df$hypertension)
stroke_df$heart_disease <- as.factor(stroke_df$heart_disease)
stroke_df$ever_married <- as.factor(stroke_df$ever_married)
stroke_df$bmi <- as.numeric(stroke_df$bmi)
stroke_df$stroke<- as.factor(stroke_df$stroke)
str(stroke_df)

# Finding if there are any NA's in the dataset
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
summary(stroke_df)

```
```{r}
# There are 201 NA's in BMI. Replacing those variables with mean of BMI
stroke_df$bmi[is.na(stroke_df$bmi)] <- mean(stroke_df$bmi,na.rm = TRUE)
# Finding if there are any NA's after imputation of null values
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
```

```{r}

df<-table(stroke_df$stroke)
barplot(df,main='Distribution of Target Variable (Stroke)',col=c('green','red'))
```


So from the plot, we can see that the number of observations of 0 is greater than the number of observations 1. From this we can see that the dataset is imbalanced.

```{r}
dat=table(stroke_df$stroke,stroke_df$smoking_status)
barplot(dat,
main = "People segregated based on their smoking habbits",
xlab = "Smoking status",
ylab = "Number of people",
col = c("Green","red"),
beside = TRUE
)
```
