---
title: "Brain Stroke Prediction - Code & Technical Analysis"
author: "Team Sage"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: TRUE
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r basic_libraries, include=FALSE}
library(ggplot2)
library(gridExtra)
library(corrplot)
library(ROSE)
library(dplyr)
```
# Loading the dataset
```{r load_csv, include=TRUE, echo=TRUE}
# reading the dataset
stroke_df = read.csv("healthcare-dataset-stroke-data.csv")
```

```{r desc, include=TRUE, echo=TRUE}
# structure of the dataset
str(stroke_df)
```
# Data Cleaning and summary of the dataset

```{r data_cleaning, include=TRUE, echo=TRUE}
stroke_df = subset(stroke_df, select = -c(id))

# converting the numeric variables to factor variables
stroke_df$gender = as.factor(stroke_df$gender)
stroke_df$hypertension = as.factor(stroke_df$hypertension)
stroke_df$heart_disease = as.factor(stroke_df$heart_disease)
stroke_df$ever_married = as.factor(stroke_df$ever_married)
stroke_df$bmi = as.numeric(stroke_df$bmi)
stroke_df$stroke = as.factor(stroke_df$stroke)

# To get the summary statistics of the dataset
summary(stroke_df)

```

```{r find_na_fix, include=TRUE, echo=TRUE}
# To find the NA's in the dataset
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))

```

```{r na_fix, include=TRUE, echo=TRUE}

# Replacing NA values with average BMI value
stroke_df$bmi[is.na(stroke_df$bmi)] = mean(stroke_df$bmi,na.rm = TRUE)
paste("The NA's in the dataset after imputation of BMI with mean is:",sum(is.na(stroke_df)))
```

```{r summary_after_cleaning, include=TRUE, echo=TRUE}
# since we had only 1 data point in others category for gender we are removing it
stroke_df = subset(stroke_df,gender!="Other")
# summary of the dataset after cleaning 
summary(stroke_df)
```
# Subsetting of the dataset

```{r subsets, include=TRUE, echo=TRUE}
# subsetting the data for various analyses 
stroke_1 = subset(stroke_df, stroke == 1)
stroke_0 = subset(stroke_df, stroke == 0)
stroke_1_female = subset(stroke_df, stroke == 1 & gender == "Female")
stroke_1_male = subset(stroke_df, stroke == 1 & gender == "Male")

# creating different columns for bmi, age and average_glucose_level based on different bucketing for each variables
dat <- within(stroke_df, {   
  bmi.cat = NA # need to initialize variable
  bmi.cat[bmi < 18.5] = "underweight"
  bmi.cat[bmi >= 18.5 & bmi < 25] = "normal"
  bmi.cat[bmi >= 25 & bmi < 30] = "overweight"
  bmi.cat[bmi >= 30 & bmi < 40] = "obesity"
  bmi.cat[bmi >=40] = "severe obesity" 
  avg_gluc.cat = NA # need to initialize variable
  avg_gluc.cat[avg_glucose_level < 60] = "Below 60"
  avg_gluc.cat[avg_glucose_level >= 60 & avg_glucose_level < 90] = "60 - 90"
  avg_gluc.cat[avg_glucose_level >= 90 & avg_glucose_level < 120] = "90 - 120"
  avg_gluc.cat[avg_glucose_level >= 120 & avg_glucose_level < 180] = "120 - 180"
  avg_gluc.cat[avg_glucose_level >= 180 & avg_glucose_level < 273] = "180 - 273"
  age.cat = NA
  age.cat[age<=20] = "Under 20"
  age.cat[age>=21 & age<= 40 ] = "20-40"
  age.cat[age>=41 & age<= 60 ] = "40-60"
  age.cat[age>=61 & age<= 80 ] = "60-80"
  age.cat[age>=80 ] = "above 80"
   } )

dat$bmi.cat<- factor(dat$bmi.cat,levels= c("underweight","normal","overweight","obesity","severe obesity" ))
dat$avg_gluc.cat<- factor(dat$avg_gluc.cat,levels= c("Below 60","60 - 90","90 - 120","120 - 180","180 - 273" ))
dat$age.cat<- factor(dat$age.cat,levels= c("Under 20","20-40","40-60","60-80","above 80"))
```

# Exploratory Data Analysis

## Distribution of the variables

```{r variable_dist, include=TRUE, echo=TRUE}
# creating density plots to see the distribution of the numerical variables

ggplot(stroke_df, aes(x=age)) +  geom_density(fill="skyblue", color="skyblue", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Distribution for age")

ggplot(stroke_df, aes(x=bmi)) +  geom_density(fill="pink", color="pink", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Distribution for BMI")

ggplot(stroke_df, aes(x=avg_glucose_level)) +  geom_density(fill="lightgreen", color="lightgreen", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Distribution for Average Glucose Level")

# Count of Males and Females in the Dataset
ggplot(stroke_df, aes(x=gender,fill=gender))+geom_bar()+ggtitle("Count of Male and Female in the Dataset")+ theme_bw() +
  theme()+ xlab("Gender") + ylab("Count of people")
```

## Distribution of the target variable 

```{r target_var, include=TRUE, echo=TRUE}
ggplot(stroke_df, aes(x=stroke,fill=stroke))+geom_bar()+ggtitle("Distribution of Target variable (Stroke)")+ theme_bw() + theme()+ xlab("Stroke") + ylab("Count of people")
```

## Does Work type, BMI and Smoking status cause stroke?

```{r work_type, include=TRUE}

# Seeing if work type affects people
ggplot(stroke_df,aes(x=work_type, fill=stroke))+geom_bar(position = "dodge")+ggtitle("Work Type Vs Stroke")+ theme_bw() +
  theme()+ xlab("Work Type") + ylab("Count of people who had a stroke")
# Seeing if Smoking status causes stroke
ggplot(stroke_df,aes(x=smoking_status, fill=stroke))+geom_bar(position = "dodge")+ggtitle("Smoking status Vs Stroke")+ theme_bw() + theme()+ xlab("Smoking Status") + ylab("Count of people who had a stroke")
# BMI vs stroke 
ggplot(stroke_df, aes(x=bmi, fill=stroke)) + geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity')+ theme_bw() + theme()+ggtitle("BMI vs Stroke")
```

## Distribution of people who suffered stroke due to their smoking habit and work type 

```{r job_smoking, include=TRUE, echo=TRUE}
# work type and smoking habits does it cause stroke?
ggplot(stroke_1,aes(x=work_type, fill= smoking_status))+geom_bar(position = "dodge")+ggtitle("Distribution of people who had a stroke based on their smoking habits")+ theme_bw() + theme()+ xlab("Work type") + ylab("Count of people ")
```

The above graph only includes those who have had a stroke, and we can see from the graph that most people in the private sector and self-employed have had a stroke, and it's worth noting that the two major causes of stroke were those who had never smoked and those who had formerly smoked.

### Does it depend based on the gender?

```{r gender_job_smoking, include=TRUE, echo=TRUE}
# seeing if gender has any affect on stroke based on smoking and work type
ggplot(stroke_1_female,aes(x=work_type, fill= smoking_status))+geom_bar(position = "dodge")+ggtitle("Distribution of females who had a stroke based on their smoking habits")+ theme_bw() + theme()+ xlab("Work type") + ylab("Count of people ")

ggplot(stroke_1_male,aes(x=work_type, fill= smoking_status))+geom_bar(position = "dodge")+ggtitle("Distribution of males who had a stroke based on their smoking habits")+ theme_bw() + theme()+ xlab("Work type") + ylab("Count of people ")
```

As seen in the graph above, the key causes that caused a stroke in females were never smoked and formerly smoked, whereas in males, the reason for stroke was formerly smoked and smokes in the private sector, and formerly smoked and never smoked in the self-employed sector.

To see if these variables are dependent we did a chi square test.

```{r chi_test, include=TRUE, echo=TRUE}
ctest1 = chisq.test(dat$smoking_status, dat$stroke)
ctest2 = chisq.test(dat$work_type, dat$stroke)
ctest1
ctest2
```

As we can see from the results, the p values for both tests are less than 0.05, thus we can fail to accept the null hypothesis and infer that smoking status and work type are related to stroke.

## Does age have any affect on stroke?

```{r age_stroke_1, include=TRUE, echo=TRUE}
# seeing if age has effect in stroke
ggplot(stroke_1, aes(x = age, fill=stroke))+ geom_density(alpha = 0.3)+ ggtitle("Density plot for age of people who had stroke")+ theme_bw() + theme()
```

In general, we know that as one gets older, the likelihood of contracting an illness increases. We wanted to see if this statement held true with our dataset. The above graph clearly shows that as one's age grows, the likelihood of having a stroke increases.

We chose the ANOVA test because we wanted to examine if age played a role in a person suffering from stroke.
```{r anova_test, include=TRUE, echo=TRUE}
one.way <- aov(age ~ stroke, data = stroke_df)
summary(one.way)
```
The ANOVA test result indicates that the p value is less than 0.05 and it is very significant, implying that as age increases, the likelihood of having a stroke increases regardless of gender.

### Does BMI with age cause stroke?
```{r age_bmi, include=TRUE, echo=TRUE}

# how bmi and age is stroke in people 
ggplot(stroke_1, aes(x=bmi, y=age, color=stroke))+geom_point(size=3)+ ggtitle("Scatter plot for age vs bmi for people who had a stroke")

ggplot(stroke_0, aes(x=bmi, y=age, color=stroke))+geom_point(size=3)+ggtitle("Scatter plot for age vs bmi for people who didnt have a stroke")


ggplot(stroke_1, aes(x=bmi, y=age, color=stroke))+geom_boxplot(fill = "skyblue")+ ggtitle("Box plot for age vs bmi for people who had a stroke")


ggplot(stroke_0, aes(x=bmi, y=age, color=stroke))+geom_boxplot(fill = "skyblue")+ ggtitle("Box plot for age vs bmi for people who didnt have a stroke")

# we can see 25 % to 75 % who get stroke are from age 60 - 80 
# we can see become who have bmi from 27.5 to 32.6 have higher chances of stroke
```


The categories are encoded as:

- Underweight - BMI < 18.5
- Normal - BMI >= 18.5 and BMI <= 25
- Overweight - BMI >= 25.0 and BMI <= 30
- Obesity - BMI >= 30.0 and BMI < 40
- "Extreme" or Severe Obesity - BMI >= 40 

```{r age_bmi.cat, include=TRUE, echo=TRUE}

# boxplot to see if bmi along with age has anything to do with stroke
ggplot(dat, aes(x=bmi.cat,y = age, fill=stroke))+geom_boxplot()+ ggtitle("Boxplot for Age vs BMI(categorical)")

```



```{r age_bmi.cat_chi_test, include=TRUE, echo=TRUE}
# chi square test for bmi and stroke
ctest1 = chisq.test(dat$bmi.cat, dat$stroke)
ctest1
```
As we can see from the chi-squared test, the p-value is less than 0.05, So we fail to accept the null hypothesis and can conclude that BMI is related to stroke.

### Does age with average glucose level affect stroke?
```{r age_avg_glucose, include=TRUE, echo=TRUE}
# seeing if average glucose level with age has an effect on stroke
stroke_1_dat = subset(dat, stroke==1 )
ggplot(stroke_1_dat, aes(fill=stroke, y = age, x = avg_gluc.cat))+ geom_boxplot()+ ggtitle("Boxplot for Age vs Average Gluclose level(Categorical) for people with stroke")
```

```{r avg_glucose_level_chi_tests, include=TRUE, echo=TRUE}
# chi square test for average glucose level and stroke
ctest1 = chisq.test(dat$avg_gluc.cat, dat$stroke)
ctest1
```
We can see from this test that the p-value is less than 0.05, so we cannot accept the null hypothesis and can conclude that average glucose level is dependent on stroke.

### Does age cause hypertension?

```{r age_hypertension, include=TRUE, echo=TRUE}
# seeing if a person can get hypertension with age 
ggplot(data = stroke_df, aes(x=as.character(hypertension), y=age, fill=hypertension)) +
    geom_boxplot() +
    labs(title="Age distribution by hypertension", x="hypertension", y="age")
```

```{r cqt_hypertension, include=TRUE, echo=TRUE}
# chi square test for hypertension and stroke
ctest3 = chisq.test(stroke_df$hypertension, stroke_df$stroke)
ctest3
```
We can see from this test that the p-value is less than 0.05, so we cannot accept the null hypothesis and can conclude that hypertension is dependent on stroke.

### Does age cause heart disease?
```{r age_heart_disease, include=TRUE, echo=TRUE}
# seeing if a person can get a heart disease with age
ggplot(data = stroke_df, aes(x=as.character(heart_disease ), y=age, fill=heart_disease)) +
    geom_boxplot() +
    labs(title="Age distribution by heart_disease ", x="heart_disease ", y="age")
```

```{r cst_heart_disease, include=TRUE, echo=TRUE}
ctest4 = chisq.test(stroke_df$heart_disease, stroke_df$heart_disease)
ctest4
```
We can see from this test that the p-value is less than 0.05, so we cannot accept the null hypothesis and can conclude that heart disease is dependent on stroke.

## Does residence type cause stroke?

```{r graph_residence_type, include=TRUE, echo=TRUE}
ggplot(stroke_1_dat, aes(x=Residence_type,fill=stroke))+geom_bar(position = "dodge")+ggtitle("People Having stroke in Urban and Rural Area (People Affected by stroke)")+ theme_bw() +theme()+ xlab("Residence Type") + ylab("Number of people")
```

The above graph represents only stroke affected people in Rural and Urban Residence.

### Distribution of work type in rural and urban area for people who suffered stroke

```{r graph_residence_type_work, include=TRUE, echo=TRUE}
# seeing is the place where you and where you work cause stroke
ggplot(stroke_1_dat,aes(x=Residence_type,fill=work_type))+geom_bar(position="dodge")+ggtitle("Work Type of Rural and Urban - People who are affected by stroke ")+theme()+theme_bw()+ylab("No of people")
```

### Distribution of smoking habit in rural and urban area for people who suffered stroke
```{r graph_residence_type_smoking, include=TRUE, echo=TRUE}
# seeing if smoking habit in the place where you live cause stroke
ggplot(stroke_1_dat,aes(x=Residence_type,fill=smoking_status))+geom_bar(position="dodge")+ggtitle("Smoking habit of Rural and Urban - People who are affected by stroke ")+theme()+theme_bw()+ylab("No of people")
```

```{r cst_residence_type, include=TRUE, echo=TRUE}
# chi square test for residence type and stroke
ch<-table(stroke_df$Residence_type, stroke_df$stroke)  
chisq.test(stroke_df$Residence_type, stroke_df$stroke)
```
We can see from this test that the p-value is greater than 0.05, so we can accept the null hypothesis and can conclude that heart disease is independent on stroke.